---
title: "p8105_hw2"
author: "Liner Ge"
date: "2020/9/27"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```
# Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwhell_df=
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="Mr. Trash Wheel",
    range=cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls= round(sports_balls),
    sports_balls= as.integer(sports_balls)
  )

```

Read precipitation data for 2017 and 2018

```{r}
precip_2018=
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="2018 Precipitation",
    skip=1,
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year=2018) %>% 
  relocate(year)

precip_2017=
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="2017 Precipitation",
    skip=1,
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year=2017) %>% 
  relocate(year)
```

Combine annual precipitation.

```{r}
precip_df=
  bind_rows(precip_2017,precip_2018) %>% 
  mutate(month=month.name[month])
```

This dataset contains information from the Mr. Trashwhell collector in Baltimore, Maryland. As trash enters the harbor, the trashwhell collect the trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwhell_df)` rows in our final dataset. Additional data sheet include month precipitation data. There are a total of `r nrow(precip_df)` rows in month precipitation dataset.The total precipitation in 2018 is `r sum(pull(precip_2018,total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(select(filter(trashwhell_df, year == 2017),sports_balls),sports_balls))`.

# Problem 2

Read NYC Transit data.
```{r}
transit_df=
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry= recode(entry, YES= TRUE, NO= FALSE))
```
This is a NYC transit dataset. It contains following variables:`r names(transit_df)`. My data cleaning steps:1) I use janitor::clean_names to clean up variables names and convert them to lower snake case.2) I select important variables and remove others. The dimension of this dataset is `r nrow(transit_df)`×`r ncol(transit_df)`. I think the data is not tidy for following reasons:

 (1) Route number and route name are not distinct;
 
 (2) Value types in route1-route11 are not consistent.

Answer the following questions:

(1) There are `r nrow(distinct(transit_df,line,station_name))`  distinct stations.

(2) `r filter(transit_df, ada=="TRUE") %>% distinct(line,station_name) %>% nrow()` distinct stations are ADA compliant.

(3) The proportion of distinct station entrances / exits without vending allow entrance is `r nrow(filter(transit_df, entry=="TRUE" & vending=="NO"))/ nrow(filter(transit_df, vending== "NO"))` .

```{r}
reformat_data = 
  transit_df %>% 
  mutate_at(vars(route1:route11),as.factor) %>% 
  pivot_longer(
    route1:route11,
    names_to= "route_name",
    names_prefix = "route",
    values_to ="route_number") %>% 
  drop_na(route_number)
  
```

(1) `r filter(reformat_data, route_number=="A") %>% distinct(line,station_name) %>% nrow()` distinct stations serve the A train.

(2) Of the stations that serve the A train, `r filter(reformat_data, route_number=="A"& ada=="TRUE") %>% distinct(line,station_name) %>% nrow()` are ADA compliant.

# Problem 3

## First
```{r}
pols_month=
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into=c("year", "month", "day"),sep="-") %>% 
  mutate_at(vars(month),as.numeric) %>% 
  mutate(month=month.name[month]) %>% 
  mutate(president=ifelse(prez_gop >=1,"gop","dem")) %>% 
  select(-prez_dem,-prez_gop,-day)
  
         
```

## Second
```{r}
snp=
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(date, into=c( "month", "day", "year"),sep="/") %>% 
  mutate_at(vars(month),as.numeric) %>% 
  arrange(year,month) %>% 
  mutate(month=month.name[month]) %>% 
  relocate(year,month)%>% 
  select(-day)
 
```

## Third
```{r}
unemployment=
  read.csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to= "month",
    values_to ="percentage") %>% 
  mutate(month=recode(month,"jan"="January","feb"="February","mar"="March","apr"="April","may"="May","jun"="June","jul"="July","aug"="August","sep"="September","oct"="October","nov"="November","dec"="December"))%>% 
  mutate_at(vars(year),as.character)
```

Join the datasets.
```{r}

pols_snp=left_join(pols_month, snp) 
final_data=left_join(pols_snp,unemployment)
```

The dataset pols_month is a `r nrow(pols_month)` × `r ncol(pols_month)` dataset. It contains variables: `r names(pols_month)`. The range of year is `r range(pull(pols_month,year))`.

The dataset snp is a `r nrow(snp)` × `r ncol(snp)` dataset. It contains variables: `r names(snp)`. The range of year is `r range(pull(snp,year))`.

The dataset unemployment is a `r nrow(unemployment)` × `r ncol(unemployment)` dataset. It contains variables: `r names(unemployment)`. The range of year is `r range(pull(unemployment,year))`.

The dataset final_data is a `r nrow(final_data)` × `r ncol(final_data)` dataset. It contains variables: `r names(final_data)`. The range of year is `r range(pull(final_data,year))`.
