---
title: "p8105_hw2"
author: "Liner Ge"
date: "2020/9/27"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```
# Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwhell_df=
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="Mr. Trash Wheel",
    range=cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls= round(sports_balls),
    sports_balls= as.integer(sports_balls)
  )

```

Read precipitation data for 2017 and 2018

```{r}
precip_2018=
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="2018 Precipitation",
    skip=1,
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year=2018) %>% 
  relocate(year)

precip_2017=
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet="2017 Precipitation",
    skip=1,
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year=2017) %>% 
  relocate(year)
```

Combine annual precipitation.

```{r}
precip_df=
  bind_rows(precip_2017,precip_2018) %>% 
  mutate(month=month.name[month])
```

This dataset contains information from the Mr. Trashwhell collector in Baltimore, Maryland. As trash enters the harbor, the trashwhell collect the trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwhell_df)` rows in our final dataset. Additional data sheet include month precipitation data. There are a total of `r nrow(precip_df)` rows in month precipitation dataset.The total precipitation in 2018 is `r sum(pull(precip_2018,total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(select(filter(trashwhell_df, year == 2017),sports_balls),sports_balls))`.